Dependent Object Types

### 2012-09-05 doc/candidacy/slides.tex
starting tribe section

\begin{frame}{Translucent Sums}
\begin{itemize}
\item values representing modules
\item translucent sum: sequence of bindings; translucent sum type:
  sequence of declarations
\item differs from traditional records, because later fields can
  depend on earlier ones
\item allows any type to be partially or fully deermined by a type
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Example}
\begin{minted}{sml}
structure XInt = struct
  type T = int
  val v = 3
  val f = negate
end
structure XBool = struct
  type T = bool
  val v = true
  val f = not
end
structure X =
  if flip() then XInt else XBool
signature XType = sig
  type T
  val v = T
  val f = T -> T
end
\end{minted}
\end{frame}

@@

\begin{frame}{Path-Dependent Types}
\begin{itemize}
\item Scala's abstract types have close resemblances to abstract types
  of signatures in these ML-like module systems. In particular, in
  their calculus, there's a form of dependent type, since translucent
  sums are terms that may contain types.
\item More concretely, their so-called VALUE rules enable a
  translucent sum to be given a more precise type by referring to the
  names of its type members with a path selection -- this ability
  based on path-dependent types is not only essential for breaking the
  dependencies between (sub)fields but also critical for the
  propagation of typing information.
\item Compared to Scala, these path-dependent types are more limited
  because recursive references and bounded quantifications are not
  allowed.
\end{itemize}
\end{frame}

\subsection{Tribe Calculus}
% Virtual Classes
% Example
% Tribe Types (generalized paths & subtyping relation)
% Follow-up
  % Ownership System?
  % Soundness?
  % Cross-Family Inheritance

>>>>

\begin{frame}{ML: Translucent Sums}
\begin{itemize}
\item values representing modules
\item translucent sum: sequence of bindings; translucent sum type:
  sequence of declarations
\item differs from traditional records, because later fields can
  depend on earlier ones
\item allows any type to be partially or fully deermined by a type
\end{itemize}
\end{frame}

\begin{frame}[fragile]{ML: Example}
\begin{minted}{sml}
structure XInt = struct
  type T = int
  val v = 3
  val f = negate
end
structure XBool = struct
  type T = bool
  val v = true
  val f = not
end
structure X =
  if flip() then XInt else XBool
signature XType = sig
  type T
  val v = T
  val f = T -> T
end
\end{minted}
\end{frame}

@@

\begin{frame}{ML: Path-Dependent Types}
\begin{itemize}
\item Scala's abstract types have close resemblances to abstract types
  of signatures in these ML-like module systems. In particular, in
  their calculus, there's a form of dependent type, since translucent
  sums are terms that may contain types.
\item More concretely, their so-called VALUE rules enable a
  translucent sum to be given a more precise type by referring to the
  names of its type members with a path selection -- this ability
  based on path-dependent types is not only essential for breaking the
  dependencies between (sub)fields but also critical for the
  propagation of typing information.
\item Compared to Scala, these path-dependent types are more limited
  because recursive references and bounded quantifications are not
  allowed.
\end{itemize}
\end{frame}

\subsection{Tribe Calculus}

\begin{frame}{Tribe Calculus: Virtual Classes}
\begin{itemize}
\item {\it familes} of classes inherited together
\item classes are lexically nested inside other classes: when a class
  is inherited, its nested inner classes are inherited along with
  their methods and fields
\item two form of inheritance: subclassing and further binding
\item {\it family polymorphism}: code written for one family
  also works for extensions of that family
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Tribe Types}
\begin{itemize}
\item In Tribe, types are a form of generalized paths: a final
  variable, {\tt this} or a class name then followed by a possibly
  empty sequence of field, {\tt out} or class selections.
\item {\tt kitt.Passenger.name}: the name of one of Kitt's passengers
\item {\tt kitt.driver <: Car.driver <: Car.Passenger <: Car.Traveller <:
  Vehicle.Traveller}.
\item {\tt kitt.driver $\not<:$ karr.driver} and {\tt kitt.Passenger $\not<:$ karr.Passenger}.
\end{itemize}
\begin{minted}[frame=lines,fontsize=\footnotesize]{java}
class Vehicle { class Traveller { ... } }
class Car extends Vehicle {
  class Passenger extends Traveller {
    class String { ... }
    final String name;
  }
  final Passenger driver;
}
final Car kitt;
final Car karr;
\end{minted}
\end{frame}

% Virtual Classes
% Example
% Tribe Types (generalized paths & subtyping relation)
% Follow-up
  % Ownership System?
  % Soundness?
  % Cross-Family Inheritance
